---
title: "DATA 300 3 Homework 6 Solution"
author: "Aadarsha Gopala Reddy"
date: "December 7, 2022"
output:
  pdf_document:
    toc: true
  html_document:
    toc: true
    toc_float: true
    df_print: paged
---

# 0. Loading the libraries

```{r import-libraries}
# load the libraries
library(tidyverse)
library(tidytext)
library(tidyr)
```

---

# 1. Setup

Load the packages "keras" and "tensorflow." If this is your first time using these packages, you may have to install the underlying python libraries using the following functions (these functions do not take any arguments):

- install_tensorflow()
- install_keras()


Note: Windows machines may experience an OpenSSL error when attempting to install these packages. This is caused by a bug in the reticulate package, which is automatically loaded when using python libraries. You can work around this problem by installing a developer version of the reticulate package, which has implemented a bug fix:

`remotes::install_github("rstudio/reticulate")`

```{r setup}
library(tensorflow)
install_tensorflow()
library(keras)
install_keras()
```

---

# 2. Load the Dataset

Load the Fasion-MNEST dataset using the function `dataset_fashion_mnest()`. Store this dataset as an object in R. (Unlike datasets we have seen in the past, this dataset is a list containing multiple sub-objects.)

```{r load-the-dataset}
# load the dataset
fashion_mnist <- dataset_fashion_mnist()
```

---

# 3. Format the Dataset

Using the dataset that you loaded in the previous question, create four objects in R:

- Image data for the training dataset;
- Class labels for the training dataset;
- Image data for the testing dataset;
- Class labels for the testing dataset.

Hint: use the dollar sign operator to select individual components of the Fashion-MNEST dataset.

```{r format-the-dataset}
# format the dataset
train_images <- fashion_mnist$train$x
train_labels <- fashion_mnist$train$y
test_images <- fashion_mnist$test$x
test_labels <- fashion_mnist$test$y
```

---

# 4. Recode Class Labels

The class labels for both the testing and training dataset are stored as numeric values 0 through 9. Recode the class labels into two categories: one for items worn on the torso, and another for pants/shoes/accessories. In your new coding, T-shirts/tops, pullovers, dresses, coats, and shirts will be coded 1, and all other classes will be coded 0.

When you are finished, you should have two objects:

- Training class labels, containing 60,000 instances of either 0 or 1;
- Testing class labels, containing 10,000 instances of either 0 or 1.

```{r recode-class-labels}
# recode the class labels
train_labels <- ifelse(train_labels %in% c(0, 1, 2, 3, 4, 5), 1, 0)
test_labels <- ifelse(test_labels %in% c(0, 1, 2, 3, 4, 5), 1, 0)
```

---

# 5. Pre-Process Pixel Values

The pixel brightness values are stored as numeric values between 0 and 255. Rescale these values to be bounded between 0 and 1. (This will make the computations easier for our neural net.) Values of 255 in the original coding will become values of 1 in the new coding; values of 100 in the old coding will become values of .392 in the new coding; values of 0 in the original coding will become values of 0 in the new coding; and so on.

Hint: you can accomplish this by dividing every value by 255.

```{r pre-process-pixel-values}
# pre-process the pixel values
train_images <- train_images / 255
test_images <- test_images / 255
```

---

# 6. Structure a Neural Network

Using the keras package, you can build your network structure layer-by-layer sequentially. Begin by initializing a sequential neural network:

`net <- keras_model_sequential()`

Add layers to your model using a pipe. The code below adds the first layer (a convolutional layer) to your network:

`net %>% layer_conv_2d(input_shape = c(28, 28, 1), filters = 1, kernel_size = c(15, 15))`

Add additional layers in order by adding more pipes:

`net %>% [layer1] %>% [layer2] %>%` ...and so on.

Check the layers you have added to your network by typing its name. R will output a line for each layer, the output shape of each layer, and some summary information about the model.

Create a network structure containing the following layers:

- A convolutional layer
  + Arguments:
    * This layer should take an input of a 2-dimensional array of 28x28 pixels;
    * This layer should apply a single convolutional filter to the input;
    * The convolutional filter should be 15x15 pixels;
  + Outputs:
    * This layer should output a 14x14 output feature map.
- A max-pooling layer
  + Arguments:
    * This layer should "pool" values by taking the maximum value within each 2x2 window.
  + Outputs:
    * This layer should output a 7x7 pooled feature map.
- A flattening layer
  + Arguments:
    * This layer should take, as an input, the 7x7 pooled feature map produced by the previous layer.
  + Output:
    * This layer should output a vector of 49 values; these values should be "flattened" in the sense that they are no longer structured as a 2-dimensional array, but simply as a vector of 49 individual numeric values.
- A dense (AKA "fully-connected" or "hidden") layer
  + Arguments:
    * This layer should contain 49 nodes (one for each value it will take as an input);
    * This layer should process these values using the "relu" activation function. (The ReLU, or rectified linear unit, function is a popular function in neural networks; it changes negative values to 0, which speeds up computation time and introduces nonlinearity into the network.)
  + Output:
    * This layer should output a vector of 49 numeric values (one for each node).
- An output layer: a dense layer that generates an output value
  + Arguments:
    * This layer should contain one single node;
    * This layer should use a "sigmoid" activation function to transform its inputs into a single value between 0 and 1.
  + Output:
    * This layer should output a single value bounded between 0 and 1; this is the predicted probability that an image depicts a top.

```{r structure-a-neural-network}
# structure the neural network
net <- keras_model_sequential() %>%
  layer_conv_2d(input_shape = c(28, 28, 1), filters = 1, kernel_size = c(15, 15)) %>%
  layer_max_pooling_2d(pool_size = c(2, 2)) %>%
  layer_flatten() %>%
  layer_dense(units = 49, activation = "relu") %>%
  layer_dense(units = 1, activation = "sigmoid")
```

---

# 7. Compile the Network

Compile the network that you created in the previous step.

`net %>% compile(...)`

The compile function should take the following arguments:

- The compiler should use the "adam" optimizer;
- The compiler should use the "binary_crossentropy" loss function;
- The compiler should use the "accuracy" metric.


```{r compile-the-network}
# compile the network
net %>% compile(optimizer = "adam", loss = "binary_crossentropy", metrics = "accuracy")
```

---

# 8. Train the Network

Train your neural network on the images and class labels for the training dataset. 

`net %>% fit(...)`

```{r train-the-network}
# train the network
net %>% fit(train_images, train_labels, epochs = 10)
```

# 9. Test the Network

Find the accuracy of your neural network in both the testing and the training data.

`net %>% evaluate(...)`

What percentage of the training images does your neural network classify correctly? What percentage of the testing images does your neural network classify correctly?

```{r test-the-network}
# test the network
net %>% evaluate(test_images, test_labels)
```

---