---
title: "DATA 300 3 Homework 1 Solution"
author: "Aadarsha Gopala Reddy"
date: "September 8, 2022"
output:
  pdf_document:
    toc: true
  html_document:
    toc: true
    toc_float: true
    df_print: paged
---

# 1. Load Packages

In a chunk of code, load the tidyverse package and any other packages you will use in this document.

>`install.packages()` Installs any packages.

>`library()` Loads any packages.

``` {r load-packages}
library("tidyverse")
```

# 2. Import Data

Import *protest_data.csv* as an object in R. Print the first few rows of the dataset using the `head()` function.

>`read_csv()` reads a csv file into R.

>`head()` prints the first few rows of a dataset.

``` {r import-data}
protest_data <- read_csv("protest_data.csv")
head(protest_data)
```

# 3. Filter Data

Subset the protest dataset to only the protests that occurred in the year 2022. Make sure to save your changes to the dataset.

>`filter()` filters a dataset based on a condition. This was used to filter the dataset to include only the protests that occurred in 2022.

> `%>%` pipes the output of one function into the input of another function. `tally()` counts the number of rows in a dataset. Both were used together to calculate the number of protests that occurred in 2022.

``` {r filter-data}
protest_data <- filter(protest_data, year == 2022)
head(protest_data)
protest_data %>% tally()
```

How many protests occurred in the year 2022? **8305 protests.**

# 4. Create a New Variable

Create a new variable in the protest dataset called `cont_us`. This binary variable should measure whether the protest occurred in the contiguous United States.
- Protests that occurred in Hawaii or Alaska should have a value of 0.
- Protests that occurred in the other 48 states and Washington D.C. should have a value of 1.

>`if_else()` creates a new variable based on a condition. This was used to create the `cont_us` variable.

>`select()` selects a subset of columns from a dataset. This was used to select the `state` and `cont_us` column from the dataset.

>`filter()` and `%>%` was used to calculate the number of protests that occurred outside of the contiguous United States.

``` {r create-a-new-variable}
protest_data$cont_us <- if_else(protest_data$admin1 %in% c("Hawaii", "Alaska"), 0, 1)
select(protest_data, "admin1", "cont_us")
filter(protest_data, cont_us == 0)  %>% tally()
```

How many protests occurred outside of the contiguous United States? **66 protests occurred outside the contiguous USA.**

# 5. Transform the Dataset

Transform the protest-level dataset into a state-level dataset and save it as a new object in R. In the new dataset, each observation should be a state or territory of the United States. It should contain the following variables:

- state: the name of the state.
- fatalities: the total number of fatalities at protests in that state in 2022.
- count: a count of the number of protests in that state in 2022.

Print the first few rows of the new dataset using the `head()` function.

>`group_by()` groups a dataset by a variable. This was used to group the dataset by state.

>`summarize()` summarizes a dataset by a variable. This was used to calculate the total number of fatalities and the number of protests in each state.

>`%>%` was used to pipe the output of `group_by()` into the input of `summarize()`.

``` {r transform-the-dataset}
state_data <- protest_data %>% group_by(admin1) %>% summarise(fatalities = sum(fatalities), count = n())
head(state_data)
```

# 6. Sort the State-Level Data

Sort the state-level dataset by the number of protests occurring in the state. Using a method of your choosing, display the names and number of protests for the 5 states with the most protests in 2022.

>`arrange()` arranges a dataset by a variable. This was used to arrange the dataset by the number of protests in each state.

>`select()` was used to select the `state` and `count` columns from the dataset.

>`%>%` was again used to pipe the output of `arrange()` into the input of `select()`.

``` {r sort-the-state-level-data}
state_data <- state_data %>% arrange(desc(count)) %>% select(admin1, count)
head(state_data)
```

# 7. Merge Two Datasets

Import a second dataset, `state.csv`. Each observation in this dataset is a state. This dataset contains two variables:

- state: the name of the state.
- population: the number of people living in the state.

Join your state-level dataset together with state.csv such that the population variable appears in your state-level dataset. Use the `head()` function to print your state-level dataset with the population variable included.

>`left_join()` joins two datasets together. This was used to join the state-level dataset with the state.csv dataset.

>`select()` was used to select the `state`, `count`, and `population` columns from the dataset.

>`%>%` was again used to pipe the output of `left_join()` into the input of `select()`.

``` {r merge-two-datasets}
state_population <- read_csv("state.csv")
state_data <- left_join(state_data, state_population, by = c("admin1" = "state"))
head(state_data)
```