---
title: "Homework 1 Solution"
author: "Aadarsha Gopala Reddy"
date: "September 8, 2022"
output:
  pdf_document:
    toc: true
  html_document:
    toc: true
    toc_float: true
    df_print: paged
---

# 1. Load Packages

In a chunk of code, load the tidyverse package and any other packages you will use in this document.

`install.packages()` installs a package, if it exists.
`library()` loads the package.

``` {r load-packages}
library("tidyverse")
```

# 2. Import Data

Import *protest_data.csv* as an object in R. Print the first few rows of the dataset using the `head()` function.

``` {r import-data}
protest_data <- read_csv("HW1/protest_data.csv")
head(protest_data)
```

# 3. Filter Data

Subset the protest dataset to only the protests that occurred in the year 2022. Make sure to save your changes to the dataset.

``` {r filter-data}
protest_data <- filter(protest_data, year == 2022)
head(protest_data)
protest_data %>% tally()
```

How many protests occurred in the year 2022? **8305 protests.**

# 4. Create a New Variable

Create a new variable in the protest dataset called `cont_us`. This binary variable should measure whether the protest occurred in the contiguous United States.
- Protests that occurred in Hawaii or Alaska should have a value of 0.
- Protests that occurred in the other 48 states and Washington D.C. should have a value of 1.

``` {r create-a-new-variable}
protest_data$cont_us <- if_else(protest_data$admin1 %in% c("Hawaii", "Alaska"), 0, 1)
select(protest_data, "admin1", "cont_us")
filter(protest_data, cont_us == 0)  %>% tally()
```

How many protests occurred outside of the contiguous United States? **66 protests occurred outside the contiguous USA.**

# 5. Transform the Dataset

Transform the protest-level dataset into a state-level dataset and save it as a new object in R. In the new dataset, each observation should be a state or territory of the United States. It should contain the following variables:

- state: the name of the state.
- fatalities: the total number of fatalities at protests in that state in 2022.
- count: a count of the number of protests in that state in 2022.

Print the first few rows of the new dataset using the head() function.

``` {r transform-the-dataset}
state_protest_data <- protest_data %>% add_count(admin1, wt = fatalities) %>%
  group_by(admin1) %>%
  summarize(fatalities = sum(fatalities), n=sum(n))
head(state_protest_data)
select(state_protest_data, "admin1", "fatalities", "n")
```