---
title: "Homework 1 Solution"
author: "Aadarsha Gopala Reddy"
date: "September 8, 2022"
output:
  pdf_document:
    toc: true
  html_document:
    toc: true
    toc_float: true
    df_print: paged
---

# 1. Load Packages

In a chunk of code, load the tidyverse package and any other packages you will use in this document.

`install.packages()` installs a package, if it exists.
`library()` loads the package.

``` {r load-packages}
library("tidyverse")
```

# 2. Import Data

Import protest_data.csv as an object in R. Print the first few rows of the dataset using the head() function.

``` {r import-data}
protest_data <- read_csv("protest_data.csv")
head(protest_data)
```

# 3. Filter Data

Subset the protest dataset to only the protests that occurred in the year 2022. Make sure to save your changes to the dataset.

``` {r filter-data}
protest_data <- filter(protest_data, year == 2022)
head(protest_data)
protest_data %>% tally()
```

How many protests occurred in the year 2022? **8305 protests**